Implementing a Capability Machine model into Iris

### 2020-01-15 prisc-2020-talk/slides.tex
slide fixes

@@

\column{.45\textwidth} % Left column and width
\begin{tikzpicture}
    \node<1> (img1) {\includegraphics[scale=0.7]{images/cap_1.pdf}};
    \node<2> (img2) {\includegraphics[scale=0.7]{images/cap_2.pdf}};
    \node<3> (img3) {\includegraphics[scale=0.7]{images/cap_3.pdf}};
    \node<4-> (img4) {\includegraphics[scale=0.7]{images/cap_4.pdf}};
\end{tikzpicture}

@@

\end{columns}
\end{frame}

%------------------------------------------------
%\subsection{Enforcing Local Stack Encapsulation using Capabilities} % A subsection can be created just before a set of slides with a common theme to further break down your presentation into chunks
%
%\begin{frame}[fragile]
%\frametitle{Capability Safety}
%\textbf{Local State Encapsulation}
%\\[1cm]
%\begin{columns}[c]
%
%\column{.45\textwidth} % Left column and width
%\begin{tikzpicture}
%    \node<1> (img1) {\includegraphics[scale=0.7]{images/wbcf_1.pdf}};
%    \node<2> (img2) {\includegraphics[scale=0.7]{images/wbcf_2.pdf}};
%    \node<3> (img3) {\includegraphics[scale=0.7]{images/wbcf_3.pdf}};
%\end{tikzpicture}
%
%\column{.5\textwidth} % Right column and width
%\begin{center}
%\begin{lstlisting}
%push r_stk 1
%scall r 
%pop r_stk r_1
%assert r_1 1
%halt
%\end{lstlisting}
%\end{center}
%
%
%\end{columns}
%\end{frame}

%------------------------------------------------
\subsection{Enforcing Well Bracketed Control Flow using Capabilities} % A subsection can be created just before a set of slides with a common theme to further break down your presentation into chunks

\begin{frame}[fragile]
\frametitle{Capability Safety}
\textbf{Well Bracketed Control Flow}
\\[1cm]
\begin{columns}[c]

\column{.45\textwidth} % Left column and width
\begin{tikzpicture}
    \node<1> (img1) {\includegraphics[scale=0.7]{images/wbcf_1.pdf}};
    \node<2> (img2) {\includegraphics[scale=0.7]{images/wbcf_2.pdf}};
    \node<3> (img3) {\includegraphics[scale=0.7]{images/wbcf_3.pdf}};
    \node<4> (img4) {\includegraphics[scale=0.7]{images/wbcf_4.pdf}};
    \node<5> (img5) {\includegraphics[scale=0.7]{images/wbcf_5.pdf}};
    \node<6> (img6) {\includegraphics[scale=0.7]{images/wbcf_6.pdf}};
\end{tikzpicture}

@@

\end{columns}
\end{frame}

@@

\foreach \i/\j in {1-1/5-1}
    \draw (a-\i) -- (a-\j);
%
\end{tikzpicture}
}

@@

\begin{frame}
\frametitle{Calling Convention}
\begin{tikzpicture}
    \node<1> (img1) {\includegraphics[scale=0.7]{images/calling_1.pdf}};
    \node<2> (img2) {\includegraphics[scale=0.7]{images/calling_2.pdf}};
\end{tikzpicture}
%
%\begin{longtable}[c]{|l|l|}
%\hline
%\texttt{r\_stk} & $(RWLX,Local,b,e,a)$ \\ \hline
%\endfirsthead
%%
%\endhead
%\end{longtable}

@@

\begin{itemize}
	\item<2-> using a Program Logic 
	\item<3-> using a logical relation to capture invariants on the type system
	\item<4-> \textbf{using a logical relation on an \emph{untyped} (or \emph{uni-typed}) language to capture semantic properties of the language}
\end{itemize}

@@

\textbf{Roadmap}
	\begin{itemize}
		\item<2> embed the language into Iris
		\item<3> define a program logic by proving Hoare Triples
		\item<4> define the logical relation -- using Iris tools to solve the world circularity problem
		\item<5> prove the fundamental theorem of logical relations
		\item<6> use the logical relation to prove examples that rely on local state encapsulation and well-bracketed control flow with calls to unknown adversary
	\end{itemize}

@@

%------------------------------------------------
\section{Program Logic}
%------------------------------------------------

%%------------------------------------------------
%\section{Proving Hoare Triples}
%
%%------------------------------------------------
%\subsection{Successful Execution}
%
%\begin{frame}
%\frametitle{Hoare Triples of the Program Logic: Success}
%
%\begin{align*}
%&decode(w) = \textsf{Load dst src} \\
%\wedge &\textsf{ isCorrectPC }((p_{pc},g_{pc}),b_{pc},e_{pc},a_{pc}) \\
%\wedge &\textsf{ readAllowed } p_{src} ~\wedge~  \textsf{withinBounds } (b_{src},e_{src},a_{src})\\ \\
%\{\{\{&~ \alert<2>{\textsf{PC} \mapsto_r ((p_{pc},g_{pc}),b_{pc},e_{pc},a_{pc})} * a_{pc} \mapsto_a[p_{pc}] w \\
%&* \alert<3>{dst \mapsto_r w_{dst}} * src \mapsto_r ((p_{src},g_{src}),b_{src},e_{src},a_{src}) \\
%&* a_{src} \mapsto_a[p_{src}] w_{src} \}\}\} \\
%&\textsf{Instr Executable} \\
%\{\{\{&~  \alert<2>{\textsf{PC} \mapsto_r ((p_{pc},g_{pc}),b_{pc},e_{pc},a_{pc} + 1)}
%			  * a_{pc} \mapsto_a[p_{pc}] w
%			  \\&* \alert<3>{dst \mapsto_r w_{src}}
%			  * src \mapsto_r ((p_{src},g_{src}),b_{src},e_{src},a_{src})
%			   \\&* a_{src} \mapsto_a[p_{src}] w_{src} \}\}\}
%\end{align*}
%
%\end{frame}
%
%%------------------------------------------------
%\subsection{Failed Execution}
%
%\begin{frame}
%\frametitle{Hoare Triples of the Program Logic: Failure}
%
%\begin{align*}
%&decode(w) = \textsf{Load dst src} \\
%\wedge &\textsf{ isCorrectPC }((p_{pc},g_{pc}),b_{pc},e_{pc},a_{pc}) \\
%\wedge &~\alert<2>{\neg \textsf{readAllowed } p_{src}} ~\vee~ \neg \textsf{withinBounds } (b_{src},e_{src},a_{src})\\
%\{\{\{&~\textsf{PC} \mapsto_r ((p_{pc},g_{pc}),b_{pc},e_{pc},a_{pc}) * a_{pc} \mapsto_a[p_{pc}] w \\
%&* src \mapsto_r ((p_{src},g_{src}),b_{src},e_{src},a_{src}) \}\}\} \\
%&\textsf{Instr Executable} \\
%\{\{\{&~ \alert<2>{\textsf{FailedV}}, \top \}\}\}
%\end{align*}
%
%\end{frame}

@@

@@

\begin{itemize}
	\item Embed a capability machine into Iris
	\item Define its program logic 
	\item Mechanize a unary logical relation for an untyped capability machine language
	\item Prove the fundamental theorem of logical relations
	\item Reason about examples that rely on Local Stack Encapsulation and Well-Bracketed Control Flow with calls to an unknown adversary
\end{itemize}

@@

%----------------------------------------------------------------------------------------
% Bonus Slides
%------------------------------------------------
\subsection{The Execute Condition}

>>>>

\textit{Reasoning about a Machine with Local Capabilities}: \cite{p1}

@@

\column{.45\textwidth} % Left column and width
\begin{tikzpicture}
    \node<1> (img1) {\includegraphics[scale=0.7]{images/cap_1.pdf}};
    \node<2> (img2) {\includegraphics[scale=0.7]{images/cap_2.pdf}};
    \node<3> (img3) {\includegraphics[scale=0.7]{images/cap_3.pdf}};
    \node<4-> (img4) {\includegraphics[scale=0.7]{images/cap_4'.pdf}};
\end{tikzpicture}

@@

\end{columns}
\only<2-4>{
\begin{itemize}
	\item \only<2>{Range}
			  \only<3>{Address}
			  \only<4>{Permission}
\end{itemize}}

%------------------------------------------------
%\subsection{Enforcing Local State Encapsulation using Capabilities} % A subsection can be created just before a set of slides with a common theme to further break down your presentation into chunks
%
%\begin{frame}[fragile]
%\frametitle{Capability Safety}
%\textbf{Local State Encapsulation}
%\\[1cm]
%\begin{columns}[c]
%
%\column{.45\textwidth} % Left column and width
%\begin{tikzpicture}
%    \node<1> (img1) {\includegraphics[scale=0.7]{images/wbcf_1.pdf}};
%    \node<2> (img2) {\includegraphics[scale=0.7]{images/wbcf_2.pdf}};
%    \node<3> (img3) {\includegraphics[scale=0.7]{images/wbcf_3.pdf}};
%\end{tikzpicture}
%
%\column{.5\textwidth} % Right column and width
%\begin{center}
%\begin{lstlisting}
%push r_stk 1
%scall r 
%pop r_stk r_1
%assert r_1 1
%halt
%\end{lstlisting}
%\end{center}
%
%
%\end{columns}
%\end{frame}

%------------------------------------------------
\subsection{Enforcing Well-Bracketed Control Flow using Capabilities} % A subsection can be created just before a set of slides with a common theme to further break down your presentation into chunks

\begin{frame}[fragile]
\frametitle{Capability Safety}
\textbf{Well-Bracketed Control Flow}
\\[1cm]
\begin{columns}[c]

\column{.45\textwidth} % Left column and width
\begin{tikzpicture}
    \node<1> (img1) {\includegraphics[scale=0.7]{images/wbcf_1'.pdf}};
    \node<2> (img2) {\includegraphics[scale=0.7]{images/wbcf_2'.pdf}};
    \node<3> (img3) {\includegraphics[scale=0.7]{images/wbcf_3'.pdf}};
    \node<4> (img4) {\includegraphics[scale=0.7]{images/wbcf_4'.pdf}};
    \node<5> (img5) {\includegraphics[scale=0.7]{images/wbcf_5'.pdf}};
    \node<6> (img6) {\includegraphics[scale=0.7]{images/wbcf_6'.pdf}};
\end{tikzpicture}

@@

\begin{itemize}
	\item \only<1>{We start with a stack with range b to e}
			  \only<2>{Push some local state}
			  \only<3>{Prepare adversary stack}
			  \only<4>{Adversary possesses a return capability}
			  \only<5>{Once jumped to we get back original stack - we pop the stack and assert that local state did not change}
			  \only<6>{Prepare the adversary stack for second call}
\end{itemize}
\end{frame}

@@

\foreach \i/\j in {1-1/5-1}
    \draw (a-\i) -- (a-\j);
%
\end{tikzpicture}
\begin{itemize}
	\item Local capabilities can only be stored where you have write local permission
\end{itemize}
}

@@

\begin{frame}
\frametitle{Calling Convention}
\begin{tikzpicture}
    \node<1> (img1) {\includegraphics[scale=0.7]{images/calling_1.pdf}};
    \node<2> (img2) {\includegraphics[scale=0.7]{images/calling_2.pdf}};
\end{tikzpicture}
%
%\begin{longtable}[c]{|l|l|}
%\hline
%\texttt{r\_stk} & $(RWLX,Local,b,e,a)$ \\ \hline
%\endfirsthead
%%
%\endhead
%\end{longtable}
\begin{itemize}
	\item We want the adversary to lose any temporary capabilities (such as return capabilities) upon return of a function call
\end{itemize}
\end{frame}

@@

\begin{itemize}
	\item<2-> using a program logic 
	\item<3-> using a logical relation to capture invariants on the type system
	\item<4-> \textbf{using a logical relation on an \emph{untyped} (or \emph{uni-typed}) language to capture semantic properties of the language}
\end{itemize}

@@

\textbf{Roadmap}
	\begin{itemize}
		\item<2> embed the language into Iris
		\item<3> define a program logic by proving hoare triples
		\item<4> define the logical relation -- using Iris tools to solve the world circularity problem
		\item<5> prove the fundamental theorem of logical relations
		\item<6> use the logical relation to prove examples that rely on local state encapsulation and well-bracketed control flow with calls to unknown adversary
	\end{itemize}

@@

@@

\begin{itemize}
	\item Dotted lines: private transitions
	\item Continuous lines: public transitions
\end{itemize}

@@

%\begin{itemize}
%	\item Embed a capability machine into Iris
%	\item Define its program logic 
%	\item Mechanize a unary logical relation for an untyped capability machine language
%	\item Prove the fundamental theorem of logical relations
%	\item Reason about examples that rely on Local State Encapsulation and Well-Bracketed Control Flow with calls to an unknown adversary
%\end{itemize}

\textbf{Contributions}
\begin{itemize}
	\item First mechanization of a core model of a capability machine. The mechanization includes: 
		\begin{itemize}
			\item Embedding of a capability machine language into Iris (first embedding of a machine language into Iris)
			\item Mechanized proof of the fundamental theorem of logical relations 
			\item Mechanized proof of capability safety of two non-trivial example programs 
		\end{itemize}
\end{itemize}
\textbf{Future work}
\begin{itemize}
	\item Mechanize a proof of capability safety of the awkward example
	\item Expand the mechanization with new capabilities and calling conventions that solve current shortcomings of local capabilities
\end{itemize}

@@

%----------------------------------------------------------------------------------------
% Bonus Slides
%------------------------------------------------
%------------------------------------------------
\section{Program Logic}
%------------------------------------------------

%------------------------------------------------
\section{Proving Hoare Triples}

%------------------------------------------------
\subsection{Successful Execution}

\begin{frame}
\frametitle{Hoare Triples of the Program Logic: Success}

\begin{align*}
&decode(w) = \textsf{Load dst src} \\
\wedge &\textsf{ isCorrectPC }((p_{pc},g_{pc}),b_{pc},e_{pc},a_{pc}) \\
\wedge &\textsf{ readAllowed } p_{src} ~\wedge~  \textsf{withinBounds } (b_{src},e_{src},a_{src})\\ \\
\{\{\{&~ \alert<2>{\textsf{PC} \mapsto_r ((p_{pc},g_{pc}),b_{pc},e_{pc},a_{pc})} * a_{pc} \mapsto_a[p_{pc}] w \\
&* \alert<3>{dst \mapsto_r w_{dst}} * src \mapsto_r ((p_{src},g_{src}),b_{src},e_{src},a_{src}) \\
&* a_{src} \mapsto_a[p_{src}] w_{src} \}\}\} \\
&\textsf{Instr Executable} \\
\{\{\{&~  \alert<2>{\textsf{PC} \mapsto_r ((p_{pc},g_{pc}),b_{pc},e_{pc},a_{pc} + 1)}
			  * a_{pc} \mapsto_a[p_{pc}] w
			  \\&* \alert<3>{dst \mapsto_r w_{src}}
			  * src \mapsto_r ((p_{src},g_{src}),b_{src},e_{src},a_{src})
			   \\&* a_{src} \mapsto_a[p_{src}] w_{src} \}\}\}
\end{align*}

%------------------------------------------------
\subsection{Failed Execution}

\begin{frame}
\frametitle{Hoare Triples of the Program Logic: Failure}

\begin{align*}
&decode(w) = \textsf{Load dst src} \\
\wedge &\textsf{ isCorrectPC }((p_{pc},g_{pc}),b_{pc},e_{pc},a_{pc}) \\
\wedge &~\alert<2>{\neg \textsf{readAllowed } p_{src}} ~\vee~ \neg \textsf{withinBounds } (b_{src},e_{src},a_{src})\\
\{\{\{&~\textsf{PC} \mapsto_r ((p_{pc},g_{pc}),b_{pc},e_{pc},a_{pc}) * a_{pc} \mapsto_a[p_{pc}] w \\
&* src \mapsto_r ((p_{src},g_{src}),b_{src},e_{src},a_{src}) \}\}\} \\
&\textsf{Instr Executable} \\
\{\{\{&~ \alert<2>{\textsf{FailedV}}, \top \}\}\}
\end{align*}

\subsection{The Execute Condition}
